MSDS6372Project1

### 2021-01-23 MSDS6372Project1.Rmd
update with Forward & Backward methods. Problem with test validation levels...

>>>>

```{r}
library(tidyverse)
library(ggplot2)
library(GGally)
# You need to adjust the work environment to run this code
setwd("/Users/mingyang/Desktop/SMU/Applied Statistics/MSDS6372Project1")
# Directly import cleaned dataset and start from here
CarData = read.csv("CleanedCarData.csv")
# Convert factor variables
variables.to.factor = c("Make","Model","EngineFuelType","TransmissionType","DrivenWheels","MarketCategory","VehicleSize","VehicleStyle","Crossover","Diesel","Exotic","Luxury","HighPerformance","FactoryTuner","Performance","FlexFuel","Hatchback","Hybrid","N_A")
CarData[variables.to.factor] = lapply(CarData[variables.to.factor],factor)
summary(CarData)
# Plot agsint numeric variables
CarData %>% select(MSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity) %>% ggpairs()
# Try Log transformation on MSRP
CarData$LogMSRP = log(CarData$MSRP)
# Recheck the logMSRP against numeric variables
# LogMSRP seem to have improved linear relationship between Year and EngineHP correlations
CarData %>% select(LogMSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity) %>% ggpairs()
# Look at Popularity
boxplot(CarData$Popularity)
CarData %>% select(LogMSRP,Popularity)%>% ggpairs()
#Try LogPopularity
CarData$LogPopularity = log(CarData$Popularity)
CarData %>% select(LogMSRP,LogPopularity)%>% ggpairs()
# Look at original MSRP vs Popularity
CarData %>% select(MSRP,Popularity)%>% ggpairs()
#LogMSRP and LogPopularity seem to have improved the linearity relationship between the two
# Look at histogram of LogMSRP
CarData %>% ggplot(aes(x=LogMSRP))+geom_histogram(bins=100)
# Drop MSRP and Popularity from the original Dataset
CarData<-CarData[,-c(15,16)]
# Drop MarketCategory - 72 levels are not useful for prediction
CarData <-CarData[,-10]
str(CarData)
#Split data into 80% Train 10% Test 10% Validation set
set.seed(111)
train.test.index<-sample(1:dim(CarData)[1],10723,replace=F)
train.test = CarData[train.test.index,]
validate = CarData[-train.test.index,]
set.seed(112)
train.index = sample(1:dim(train.test)[1],9531,replace=F)
train = train.test[train.index,]
test = train.test[-train.index,]

# Use Forward variable selection method to select possible variables as predictors
library(leaps)
reg.fwd=regsubsets(LogMSRP~.,data=train,method="forward",nvmax=955)
par(mfrow=c(1,3))
bics<-summary(reg.fwd)$bic
plot(1:956,bics,type="l",ylab="BIC",xlab="# of predictors")
index<-which(bics==min(bics))
points(index,bics[index],col="red",pch=10)

adjr2<-summary(reg.fwd)$adjr2
plot(1:956,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
index<-which(adjr2==max(adjr2))
points(index,adjr2[index],col="red",pch=10)

rss<-summary(reg.fwd)$rss
plot(1:956,rss,type="l",ylab="train RSS",xlab="# of predictors")
index<-which(rss==min(rss))
points(index,rss[index],col="red",pch=10)

coef(reg.fwd,676)
# By using Forward method, 676 predictors seem to generate the smallest BIC number
# This model has selected predictors below:
# Make+Model+EngineFuelType+TransmissionType+VehicleSize+VehicleStyle+HighwayMPG+CityMPG+Luxury+FactoryTuner+Performance+FlexFuel+Hybrid+N_A+Exotic+Hatchback+LogPopularity
forward.regression = lm(LogMSRP~Make+Model+EngineFuelType+TransmissionType+VehicleSize+VehicleStyle+HighwayMPG+CityMPG+Luxury+FactoryTuner+Performance+FlexFuel+Hybrid+N_A+Exotic+Hatchback+LogPopularity,data=train)
summary(forward.regression)
par(mfrow=c(2,2))
plot(forward.regression)
library(olsrr)
ols_plot_cooksd_bar(forward.regression)
ols_plot_resid_stand(forward.regression)
# After observing the Cook's D, there is observation 4659, 9435 has extremely large Cook's D
# Observation 1523 also has large Cook's D
train[4659,] # Toyota Yaris iA
exp(train[4659,25]) 
train%>% filter(Make=="Toyota") %>% select(Make,Model,Year,LogMSRP)
# Can't really make sense to eliminate observation 4659

train[9435,] # Ferrari Superamerica 2005
exp(train[9435,25]) # Exotic Car

#Look at other Exotic Cars
train%>%filter(Exotic=="Yes")%>% select(Make,Model,Year,LogMSRP,LogPopularity)
# Don't really make sense to delete observation 4659 either

# Since there are so many categorical variables, and log transformation improvement on Year and EngineHP didn't make it to the final model
# Next section we will try reimport the dataset without transformation
```

#### Objective 1 continue without Log transformation - since there are so many categorical variables

```{r}
library(tidyverse)
library(ggplot2)
library(GGally)
# You need to adjust the work environment to run this code
setwd("/Users/mingyang/Desktop/SMU/Applied Statistics/MSDS6372Project1")
# Directly import cleaned dataset and start from here
CarData = read.csv("CleanedCarData.csv")
# Convert factor variables
variables.to.factor = c("Make","Model","EngineFuelType","TransmissionType","DrivenWheels","MarketCategory","VehicleSize","VehicleStyle","Crossover","Diesel","Exotic","Luxury","HighPerformance","FactoryTuner","Performance","FlexFuel","Hatchback","Hybrid","N_A")
CarData[variables.to.factor] = lapply(CarData[variables.to.factor],factor)
summary(CarData)
# Plot agsint numeric variables
CarData %>% select(MSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity) %>% ggpairs()
CarData = CarData[,-10] #delete MarketCategory
# Try different way of sampling by Models

#Without transformation, split dataset
set.seed(111)
train.test.index<-sample(1:dim(CarData)[1],10723,replace=F)
train.test = CarData[train.test.index,]
validate = CarData[-train.test.index,]
set.seed(112)
train.index = sample(1:dim(train.test)[1],9531,replace=F)
train = train.test[train.index,]
test = train.test[-train.index,]

str(train)

# Use Forward variable selection method to select possible variables as predictors
library(leaps)
reg.fwd=regsubsets(MSRP~.,data=train,method="forward",nvmax=955)
par(mfrow=c(1,3))
bics<-summary(reg.fwd)$bic
plot(1:956,bics,type="l",ylab="BIC",xlab="# of predictors")
index<-which(bics==min(bics))
points(index,bics[index],col="red",pch=10)

adjr2<-summary(reg.fwd)$adjr2
plot(1:956,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
index<-which(adjr2==max(adjr2))
points(index,adjr2[index],col="red",pch=10)

rss<-summary(reg.fwd)$rss
plot(1:956,rss,type="l",ylab="train RSS",xlab="# of predictors")
index<-which(rss==min(rss))
points(index,rss[index],col="red",pch=10)

# how many predictors gives the best BIC?
which(bics==min(bics))
coef(reg.fwd,345)
# By looking at this model, it has selected following predictors:
# Make+Model+EngineFuelType+EngineCylinders+DrivenWheels+VehicleStyle+HighwayMPG+CityMPG+FactoryTuner+Performance+Hybrid+N_A+Exotic+Hatchback
forward.regression = lm(MSRP~Make+Model+EngineFuelType+EngineCylinders+DrivenWheels+VehicleStyle+HighwayMPG+CityMPG+FactoryTuner+Performance+Hybrid+N_A+Exotic+Hatchback,data=train)
summary(forward.regression)
# This model has higher adjusted R-squared compared to the Log transformed model
par(mfrow=c(2,2))
plot(forward.regression)
library(olsrr)
ols_plot_cooksd_bar(forward.regression)
ols_plot_resid_stand(forward.regression)
# Cook's D plot looks okay - Much better than Log Transformed Plot
#Observation 577, 230 and 7329 seem to have problem on the residual plot
train[577,]
train[230,]
train[7329,]

########################################
# Try Backward elimination model.#######
########################################
reg.bwd=regsubsets(MSRP~.,data=train,method="backward",nvmax=955)
par(mfrow=c(1,3))
bics<-summary(reg.bwd)$bic
plot(1:956,bics,type="l",ylab="BIC",xlab="# of predictors")
index<-which(bics==min(bics))
points(index,bics[index],col="red",pch=10)

adjr2<-summary(reg.bwd)$adjr2
plot(1:956,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
index<-which(adjr2==max(adjr2))
points(index,adjr2[index],col="red",pch=10)

rss<-summary(reg.bwd)$rss
plot(1:956,rss,type="l",ylab="train RSS",xlab="# of predictors")
index<-which(rss==min(rss))
points(index,rss[index],col="red",pch=10)

# how many predictors gives the best BIC?
which(bics==min(bics))
coef(reg.bwd,307)
# Predictor for backward elimination method based on lowest BIC value
# Make+Model+EngineFuelType+EngineCylinders+DrivenWheels+VehicleStyle+HighwayMPG+Performance+Exotic+Hatchback
backward.regression = lm(MSRP~Make+Model+EngineFuelType+EngineCylinders+DrivenWheels+VehicleStyle+HighwayMPG+Performance+Exotic+Hatchback,data=train)
summary(backward.regression)
# Forward select has higher adjusted R-squared compared to backward method
par(mfrow=c(2,2))
plot(backward.regression)

ols_plot_cooksd_bar(backward.regression)
ols_plot_resid_stand(backward.regression)
#Observation 577, 230 and 7329 seem to have problem on the residual plot - Same as forward selection

############################################
# Compare Forward vs. Backward on Test set #
############################################
# Test result
test.result = test$MSRP
str(test)
test = test[,-15] #delete MSRP for prediction
#Forward
forwardPred = predict(forward.regression,test)
